---
title: "yutian_analysis"
author: "Yutian Luo"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(corrplot)
```

# 2. Verify if this association holds true in this dataset, but also explore associations of all the other covariates mentioned above and draw your own conclusions.

        the initial analysis published by FiveThirtyEight showed that income inequality was the main predictor of hate crimes rates.


```{r import data}

hatecrime_data = 
  read.csv("./data/HateCrimes.csv")

```


```{r data cleaning}

hatecrime_df = 
  hatecrime_data %>% 
  janitor::clean_names() %>% 
  mutate(
    state = as.factor(state),
    unemployment = as.factor(unemployment),
    urbanization = as.factor(urbanization),
    hate_crimes_per_100k_splc = as.numeric(hate_crimes_per_100k_splc)
  )

```

# why use gini index to represent inequality
...

```{r association}

attach(hatecrime_df)

pairs(hate_crimes_per_100k_splc~gini_index)

ggplot(hatecrime_df, aes(y = hate_crimes_per_100k_splc, x = gini_index))+
  geom_point()+
  geom_smooth(method="lm", se = TRUE)+
  labs(title = "Inequality vs Hate Crimes")+
  labs(x = "Gini Index", y = "Hate Crime Rate")


```


From our initial association analysis, there is limited association between the two variables.

We furthere analyze by linear method.

```{r lm}

# linear method
fit_mod1 = 
  lm(hate_crimes_per_100k_splc~gini_index, data = hatecrime_df)
summary(fit_mod1)

# anova
anova(fit_mod1)

```

From the t test, we conclude that we are 95% confident that the inequality has a positive linear association with the hate crime rate.

This suggests that the higher inequality is, the higher rate of hate crime will occur.

# futher interpreate
...


```{r corr matrix}

hatecrime_df = 
  hatecrime_df %>% 
  drop_na() 

cor(hatecrime_df[,c(4,5,6,7,8,9)])%>%        
  corrplot(method = "number", type = "lower", diag=FALSE, bg = "black", order = "FPC")

```

From the correlation plot, we see that there is relatively higher correlation between the population's education with median houshold income. 

The other pair of moderate correlation is between inequality and population's high school degreee percentage; inequality with percent of alien; education with percent of none white; and education with inequality.

# mlr model
Our hypothesis for mlr model:
1. gini index
2. education

# We use correlation matrix to propose for a mlr model in part 4

# Exploration

```{r fit model}

mod_1 = lm(perc_non_citizen~perc_non_white, data = hatecrime_df)
summary(mod_1)

mod_inter = lm(hate_crimes_per_100k_splc~
                 gini_index +
                 perc_population_with_high_school_degree +
                 perc_non_citizen*perc_non_white,
               data = hatecrime_df)
summary(mod_inter1)

mod_2 = lm(perc_population_with_high_school_degree~
             median_household_income,
           data = hatecrime_df)
summary(mod_2)

mod_inter2 = lm(hate_crimes_per_100k_splc~
                  gini_index +
                 perc_population_with_high_school_degree+
                 perc_population_with_high_school_degree*median_household_income +
                perc_non_citizen*perc_non_white,
                data = hatecrime_df)
summary(mod_inter2)

```




















