---
title: "Final project"
author: "Minjie Bao (mb4757)"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, message = FALSE)
options(width = 150)
```

```{r}
library(tidyverse)
library(HH)
library(leaps)
library(corrplot)    
library(caret)
library(skimr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() +  theme(legend.position = "bottom"))

options(
  ggplots2.continuous.color = "viridis",
  ggplots2.continuous.fill = "viridus"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Consider interactions between variables (e.g., urbanization), and if the case, fit stratified models. 


```{r}
data_crimes = 
  read_csv("./data/HateCrimes.csv") %>% 
  janitor::clean_names() %>% 
  mutate(unemployment = recode_factor(unemployment, "high" = 1, "low" = 0),
         urbanization = recode_factor(urbanization, "high" = 1, "low" = 0),
         hate_crimes_per_100k_splc = as.numeric(hate_crimes_per_100k_splc)
         )
skimr::skim(data_crimes)
```


```{r}
fit0 = lm(hate_crimes_per_100k_splc ~ unemployment + urbanization + median_household_income + perc_population_with_high_school_degree + perc_non_citizen + gini_index + perc_non_white, data = data_crimes)
summary(fit0)
```
There are two significant variables: perc_population_with_high_school_degree and gini_index.


```{r, message=FALSE}
## graph:
theme_set(theme_classic())                                          
ggplot(data_crimes, aes(x=gini_index, y=hate_crimes_per_100k_splc, colour=factor(urbanization))) +  geom_point(size=2, alpha = 0.8) +                     
geom_smooth(method="lm", se=F, aes(group = factor(urbanization), color = factor(urbanization), alpha = 0.4)) +                                      labs(title="interaction plot 1", x="gini_index", y="hate_crimes_per_100k_splc") +
scale_color_manual(name="urbanization", labels = c("high", "low"), values = c("blue", "red")) +     
  annotate(geom="text", x=0.55, y=1.7, label="Unusual", # Add text to the plot 
color="black", size=3)

## numerical summaries:
fit1 <- lm(hate_crimes_per_100k_splc ~ gini_index*urbanization, data=data_crimes)
summary(fit1)


#stratify:
high_urban<-filter(data_crimes, urbanization=="1") 
reg_high_urban<-lm(hate_crimes_per_100k_splc ~ gini_index, data=high_urban)
summary(reg_high_urban)

low_urban<-filter(data_crimes, urbanization=="0") 
reg_low_urban<-lm(hate_crimes_per_100k_splc ~ gini_index, data=low_urban)
summary(reg_low_urban)
```

The coefficients change from positive to negative  tell us that the effect of gini_index on hate_crimes_per_100k_splc definitely differs by urbanization status!

From the graph, two regression lines for each level category of urbanization interacts with each other. Also, from the model with the interaction term, the p value of the interaction term is 0.02343 < 0.05, so the interaction term is significant.

Since the interaction term is significant, we will do stratification:

For high urbanization, the estimated coefficient of gini_index is 6.542 > 0, which means that with one unit increase in gini_index, the estimated hate_crimes_per_100k_splc level will increase by 6.542 unit. The p value is 0.0236 < 0.05, which means at the significance level of 0.05, the estimated coefficient of gini_index is significantly different than 0.

For low urbanization, the estimated coefficient of gini_index is -2.696 < 0, which means that with one unit increase in gini_index, the estimated hate_crimes_per_100k_splc level will decrease by 2.696 unit. The p value is 0.242 > 0.05, which means at the significance level of 0.05, the estimated coefficient of gini_index is not significantly different than 0.

As we can see, the slopes of high and low urbanization have different directionality, which means the associations between gini_index and hate_crimes_per_100k_splc level have opposite directions for the two groups.


```{r, message=FALSE}
## graph:
theme_set(theme_classic())                                          
ggplot(data_crimes, aes(x=gini_index, y=hate_crimes_per_100k_splc, colour=factor(unemployment))) +  geom_point(size=2, alpha = 0.8) +                     
geom_smooth(method="lm", se=F, aes(group = factor(unemployment), color = factor(unemployment), alpha = 0.4)) +                                      labs(title="interaction plot 2", x="gini_index", y="hate_crimes_per_100k_splc") +
scale_color_manual(name="unemployment", labels = c("high", "low"), values = c("blue", "red"))

## numerical summaries:
fit2 <- lm(hate_crimes_per_100k_splc ~ gini_index*unemployment, data=data_crimes)
summary(fit2)
```
Although the visualization above indicates that an interaction might exist, from the numeric summaries, the p-value of gini_index:unemployment0 is not significant(p-value = 0.08733 > 0.05). Therefore, there's no evidence of interaction of gini_index and unemployment, which means the effect of gini_index on hate_crimes_per_100k_splc is not different with respect to the levels of unemployment.


```{r, message=FALSE}
## graph:
theme_set(theme_classic())                                          
ggplot(data_crimes, aes(x=perc_population_with_high_school_degree, y=hate_crimes_per_100k_splc, colour=factor(urbanization))) +  geom_point(size=2, alpha = 0.8) +                     
geom_smooth(method="lm", se=F, aes(group = factor(urbanization), color = factor(urbanization), alpha = 0.4)) +                                      labs(title="interaction plot 3", x="perc_population_with_high_school_degree", y="hate_crimes_per_100k_splc") +
scale_color_manual(name="urbanization", labels = c("high", "low"), values = c("blue", "red"))

## numerical summaries:
fit3 <- lm(hate_crimes_per_100k_splc ~ perc_population_with_high_school_degree * urbanization, data=data_crimes)
summary(fit3)
```
From the graph, we can see that two regression lines for each level category of urbanization are parallel. Also, from the numeric summaries, the p-value of perc_population_with_high_school_degree:urbanization0 is 0.2828 > 0.05, which is not significant. Therefore, there's no interaction between perc_population_with_high_school_degree and urbanization.


```{r, message=FALSE}
## graph:
theme_set(theme_classic())                                          
ggplot(data_crimes, aes(x=perc_population_with_high_school_degree, y=hate_crimes_per_100k_splc, colour=factor(unemployment))) +  geom_point(size=2, alpha = 0.8) +                     
geom_smooth(method="lm", se=F, aes(group = factor(unemployment), color = factor(unemployment), alpha = 0.4)) +                                      labs(title="interaction plot 4", x="perc_population_with_high_school_degree", y="hate_crimes_per_100k_splc") +
scale_color_manual(name="unemployment", labels = c("high", "low"), values = c("blue", "red"))

## numerical summaries:
fit4 <- lm(hate_crimes_per_100k_splc ~ perc_population_with_high_school_degree * unemployment, data=data_crimes)
summary(fit4)
```

Although the visualization above indicates that an interaction might exist; however, from the numeric summaries, the p-value of perc_population_with_high_school_degree:unemployment0 is 0.2292 > 0.05 which is not significant. Therefore, there's no evidence of interaction of perc_population_with_high_school_degree and unemployment, which means the effect of perc_population_with_high_school_degree on hate_crimes_per_100k_splc is not different with respect to the levels of unemployment.



Counfounder:

Condition 1)
```{r}
fit5<-lm(hate_crimes_per_100k_splc ~ urbanization, data=data_crimes)
summary(fit5)
anova(fit5) # use anova to look overall!!!!!! include all the main effects without different levels
```
urbanization not associated with the outcome

Condition 2)
```{r}
fit6<-lm(gini_index ~ urbanization, data=data_crimes)
summary(fit6)
anova(fit6)
```
urbanization associated with the exposure

```{r}
fit8<-lm(hate_crimes_per_100k_splc ~ gini_index, data=data_crimes)
summary(fit8)

fit7<-lm(hate_crimes_per_100k_splc ~ gini_index + urbanization, data=data_crimes)
summary(fit7)
anova(fit7)

```
for gini_index be on the causal path way it means gini_index impacts urbanization, urbanization impact hate_crimes_per_100k_splc, in this case gini_index not on the path way.


By adjusting to urbanization the coefficient of gini_index reduced by almost 9%. Therefore, urbanization does not confoud the relationship between gini_index and hate_crimes_per_100k_splc.

The gini_index coefficient reduced from 4.0205 in SLR to 3.66771 in MLR after adjusting for urbanization (~9% reduction). 

We can conclude that urbanization not confounds the relationship b/w gini_index and hate_crimes_per_100k_splc.


